---
title: "Stat421 3/7"
author: "Cory DeSantis, Oliver Shanklin"
date: "March 14, 2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1) Give the one-step transition probability matrix P

```{r}
P <- matrix(c(0,1,0,0,0,0,
              0.5,0,(1/3),0,(1/6),0,
              0,(2/3),0,(1/3),0,0,
              0.5,0,(1/3),0,(1/6),0,
              0,0,0,0,0.5,0.5,
              0,0,0,0,0.5,0.5), nrow = 6, byrow = TRUE)
P

```

## 2) Find the communication classes and state whether each is closed or not-closed.

The communitcation classes are: \{5, 6\} (Closed) and \{1,2,3,4\} (Not Closed).

## 3) Give the period of each state

\begin{tabular}{ c c c c c c c }
State & 1 & 2 & 3 & 4 & 5 & 6 \\ 
- & - & - & - & - & - & - \\
Period & 2 & 2 & 2 & 2 & 1 & 1
\end{tabular}

## 4) Find $lim_{n\to\infty} P(n)$

```{r}

Pn <- P%*%P
for(i in seq(1:30000))
{
  Pn <- Pn%*%P
}
round(Pn, 2)


```


### Assume that X_0 = 1.

## 5) Calculate $P(X_4 = 4)$

$P(X_4 = 4) = 0$

You can never reach state 4 in 4 time steps given you started at 1.

## 6) Write a program which will simulate the DTMC for a fixed number of time steps. Run the program repeatedly to verify your answer in #5.

```{r}
end_state <- numeric()

for(j in seq(1,1000)){
  xCur <- 1
  for(i in seq(1:4))
  {
    if(xCur == 1){
      xCur <- 2
    }
    else if(xCur == 2){
      xCur <- sample(c(1,2,3,4,5,6),size = 1, replace = TRUE, prob = c(0.5,0,(1/3),0,(1/6),0))
    }
    else if (xCur == 3){
      xCur <- sample(c(1,2,3,4,5,6),size = 1, replace = TRUE, prob = c(0,(2/3),0,(1/3),0,0))
    }
    else if (xCur == 4){
      xCur <- sample(c(1,2,3,4,5,6),size = 1, replace = TRUE,prob = c(0.5,0,(1/3),0,(1/6),0))
    }
    else if (xCur == 5){
      xCur <- sample(c(1,2,3,4,5,6),size = 1, replace = TRUE, prob = c(0,0,0,0,0.5,0.5))
    }
    else if (xCur == 6){
      xCur <- sample(c(1,2,3,4,5,6),size = 1, replace = TRUE, prob = c(0,0,0,0,0.5,0.5))
    }
  
  }
  end_state[j] <- xCur
}

sum(end_state == 4)/length(end_state)
```



## 7) Let T denote the first passage time to state 6. That is, $T = min\{n > 0 : X_n = 6|X_0 = 1\}$.


```{r}

T_count <- numeric()

for(j in seq(1:10000)){

  xCur <- 1
  i <- 1

  while(TRUE){
    if(xCur == 1){
      xCur <- 2
      i<-i+1
    }
    else if(xCur == 2){
      xCur <- sample(c(1,2,3,4,5,6),size = 1, replace = TRUE, prob = c(0.5,0,(1/3),0,(1/6),0))
      i<-i+1
    }
    else if (xCur == 3){
      xCur <- sample(c(1,2,3,4,5,6),size = 1, replace = TRUE, prob = c(0,(2/3),0,(1/3),0,0))
      i<-i+1
    }
    else if (xCur == 4){
      xCur <- sample(c(1,2,3,4,5,6),size = 1, replace = TRUE,prob = c(0.5,0,(1/3),0,(1/6),0))
      i<-i+1
    }
    else if (xCur == 5){
      xCur <- sample(c(1,2,3,4,5,6),size = 1, replace = TRUE, prob = c(0,0,0,0,0.5,0.5))
      i<-i+1
    }
    else if (xCur == 6){
      #xCur <- sample(c(1,2,3,4,5,6),size = 1, replace = TRUE, prob = c(0,0,0,0,0.5,0.5))
      T_count[j] <- i
      break
    }
  
  }
}

hist(T_count)

mean(T_count)

sum(T_count > 10)/length(T_count)

```



